Depression and suicidality as credible signals of need

### 2021-05-06 Paper.rmd
- Fixed table name after we reorganized the methods. - Changed the first sentence in the discussion based on reviewer feedback. - Added a new limitations paragraph

Responses were recorded with sliders due to the fact they allow for finer grained changes than categorical scales [@KlimekEtAl2017]. Based on findings that a slider's starting position may bias results [@liu2019], we had concerns that participants would be less likely to move away from intermediate starting values than they would be in reality. For this reason, we set each T1 slider to start fully to the left (0). For the exact wording of the question and the labels on the sliders, see table 1.

@@

Consistent with costly signaling theory, costlier signals resulted in greater perceptions of need and increased the likelihood of participants helping the signaler than the less costly signals. This effect was largest in the "brother-in-law" and "romantic partner" vignettes, both of which involved claims of assault against participants' imagined daughters. Compared to these scenarios, the effect of signal type was much smaller in both the "basketball coach" and the "thwarted marriage" vignette. In the "basketball coach" vignette, part of the reason for this may have been the lack of relatedness between participants and the alleged victim. However, this was still a surprising finding due to the relatively small size of the request (i.e., recommend the coach be suspended rather than kick out family members or a romantic partner).

@@

Overall, our results on the connection between belief in need and likelihood of help are consistent with costly signaling and bargaining approaches to depression and suicide. At T1, *Belief* and *Action* were highly correlated yet distinct; while at T2, the effect of signal on both *Belief* and *Action* was to increase the magnitude of pro-signaler responses, with the effect of signal on *T2Action* being largely mediated by signal's effect on belief in need. This relationship between belief in need and likelihood of support and the importance of signal cost should be expected in scenarios when a target of a signal of need would benefit from helping if the signaler's need was genuine but face substantial net costs if it was not. Importantly, such scenarios appear to be common among modern humans and have likely been since individuals could increase their inclusive fitness through supporting kin or important cooperative partners who could return to providing benefits to supporters once their situation returned to normal [@SugiyamaSugiyama2003; @tooby1996].

@@

<!-- However, the fact that the relationships among signal cost, belief in need, and likelihood of providing support all largely fit existing theory (narrow and broad) indicates that participants paid attention to the relevant details and responded in ways that suggest some degree of external validity. For this reason, the benefits of addressing causal relationships that came through our experimental vignette approach likely outweighed the costs. -->

Another limitation is that, compared to the US sample, far more Indian participants were filtered out by our attention checks, something which could be consistent with botting, unfamiliarity with English, or low effort responses [@kennedy2020]. Assuming this corresponded with lower-quality responses among those who passed the attention checks, the weak signal effect in the thwarted marriage vignette may simply be due to added noise rather than differences in the scenario or the effectiveness of the signals compared to those in the US.

>>>>

Responses were recorded with sliders due to the fact they allow for finer grained changes than categorical scales [@KlimekEtAl2017]. Based on findings that a slider's starting position may bias results [@liu2019], we had concerns that participants would be less likely to move away from intermediate starting values than they would be in reality. For this reason, we set each T1 slider to start fully to the left (0). For the exact wording of the question and the labels on the sliders, see table 2.

@@

Consistent with the idea that depression and suicidality are credible signals of need, the two signals resulted in greater perceptions of need and increased the likelihood of participants helping the signaler relative to less costly signals. This effect was largest in the "brother-in-law" and "romantic partner" vignettes, both of which involved claims of assault against participants' imagined daughters. Compared to these scenarios, the effect of signal type was much smaller in both the "basketball coach" and the "thwarted marriage" vignette. In the "basketball coach" vignette, part of the reason for this may have been the lack of relatedness between participants and the alleged victim. However, this was still a surprising finding due to the relatively small size of the request (i.e., recommend the coach be suspended rather than kick out family members or a romantic partner).

@@

Overall, our results on the connection between belief in need and likelihood of help are consistent with credible signaling and bargaining approaches to depression and suicide. At T1, *Belief* and *Action* were highly correlated yet distinct; while at T2, the effect of signal on both *Belief* and *Action* was to increase the magnitude of pro-signaler responses, with the effect of signal on *T2Action* being largely mediated by signal's effect on belief in need. This relationship between belief in need and likelihood of support and the importance of signal cost should be expected in scenarios when a target of a signal of need would benefit from helping if the signaler's need was genuine but face substantial net costs if it was not. Importantly, such scenarios appear to be common among modern humans and have likely been since individuals could increase their inclusive fitness through supporting kin or important cooperative partners who could return to providing benefits to supporters once their situation returned to normal [@SugiyamaSugiyama2003; @tooby1996].

@@

Another limitation is that our control included a signal in the form of a verbal request, preventing comparisons to non-signaling strategies. Similarly, our design did not include vignettes with male signalers. For this reason, we have no data on the possibility of sex differences in the effectiveness of the signaling strategies examined.

In terms of data quality our largest limitation is that, compared to the US sample, far more Indian participants were filtered out by our attention checks, something which could be consistent with botting, unfamiliarity with English, or low effort responses [@kennedy2020]. Assuming this corresponded with lower-quality responses among those who passed the attention checks, the weak signal effect in the thwarted marriage vignette may simply be due to added noise rather than differences in the scenario or the effectiveness of the signals compared to those in the US.
